SODA POP: SOft solving and Defect Aware Partial Ordering Planning

### 2016-01-03 article/article.tex
New Algorithm ! Ongoing test and performance improvement

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={SODA POP : Robust online Partial Ordering Planning for real life applications},
            pdfauthor={Paper ID\#232},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{listings}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{SODA POP : Robust online Partial Ordering Planning for real life
applications}
\author{Paper ID\#232}
\date{}
\usepackage{xcolor, listings}
\usepackage{latex/aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}

@@

@@

\begin{document}
\maketitle
\begin{abstract}
As of recent years, automated planning domain has mainly focused on
performances and advances in state space planning to improve
scalability. That orientation shadows other less efficient ways of doing
like Partial Ordering Planning (POP) that has the advantage to be much
more flexible. This approach \emph{generates} plans that can be refined
easily and can be used to achieve greater resilience, repairing
capabilities and soft resolution that applications \emph{based on}
online plan recognition or decision-making can benefit from. \emph{This
paper presents} a set of algorithms, named Soft Ordering and Defect
Aware Partial Ordering Planning (SODA POP), that aims at targeting these
objectives and maintain a high plan quality. Our algorithms create
proper offline plans for goals and use an effective defect fixing
\emph{algorithm} to repair input plans even if these plans are
corrupted. SODA POP can also use healer actions and links in order to
always return a valid plan even in cases where the problem is impossible
to solve using problem derivation. Some relevant properties of these
algorithms are analyzed in this article and experimental results show
interesting performances for online planning and repairing.
\end{abstract}

@@

For some time Partial Order Planning (POP) has been the most popular
approach to planning resolution. This kind of algorithms are based on
\emph{least commitment strategy} on plan step ordering that can allow
actions to be flexibly interleaved during execution {[}1{]}. Thus, the
way the search is made using flexible partial plan as a search space
allowed for more versatility for a wide variety of uses. As of more
recent years, new state space search models and heuristics {[}2{]} have
been demonstrated to be more efficient than POP planners due to the
simplicity and relevance of states representation opposed to partial
plans {[}3{]}. This have made the search of performance the main axis of
research for planning in the community .

While this goal is justified, it shadows other problems that some
practical applications cause. For example, the flexibility of Plan Space
Planning (PSP) is an obvious advantage for applications needing online
planning: plans can be repaired on the fly as new informations and
objectives enter the system. The idea of using POP for online planning
and repairing plans instead of replanning everything is not new {[}4{]},
but has never been paired with the resilience that some other cognitive
applications may need, especially when dealing with sensors noise in
input data.

This resilience makes fixing errors easier than with an external
algorithm as the plan logic allows for context driven decision on the
way to repair the issues. For example if an action becomes unrelevant or
incoherent, the flaw in the partial plan makes the issue explicit and
therefore easier to fix. Softwares might sometimes provide plans that
can contain errors and imperfections that will decrease significantly
the efficiency of the computation and the quality of the output plan.

Adding to that, these plans may become totally unsolvable. This problem
is to our knowledge not treated in planning of all forms (state
planning, PSP, and even constraint satisfaction planning) as usually the
aim is to find a solution relative to the original plan (which makes
sense). But as we proceed a mechanism of \emph{problem derivation} may
be required. This will allow soft solving of any problem regardless of
its resolvability.

One of the applications that needs these improvements is plan
recognition with the particular use of off-the-shelf planners to infer
the pursued goal of an agent where online planning and resilience is
particularly important {[}5{]}. This method adds dummy actions that need
to be satisfied by modifying the goal to ensure that the observed plan
is selected by the planner. A system that is able to repair plan in real
time will ease such an application. Another application is
decision-making in dynamical environments. Indeed, having a plan that
details all steps and explicits the ones that are not possible, can help
decide of the plan of action to take.

These problems call for new ways to improve the answer of a planner.
These improvements must provide relevant plan information pointing out
exactly what needs to be done in order to make a planning problem
solvable, even in the case of obviously inconsistent input data. This
paper aims to solve this while preserving the advantages of PSP
algorithms (flexibility, easy fixing of plans, soundness and
completeness). Our Soft Ordering and Defect Aware Partial Ordering
Planning (SODA POP) algorithm will target those issues.

Our new set of auxiliary algorithms allows making POP algorithm more
resilient, efficient and relevant. This is done by pre-emptively
computing proper plans for goals, by solving new kinds of defects that
input plans may exhibit, and by healing compromised plan by deriving the
initial problem with forged actions to allow resolution.

To explain this work we first introduce a few notions, notations and
specificities of existing POP algorithms. Then we present and illustrate
their limitations, categorizing the different defects arising from the
resilience problem and explaining our auxiliary algorithms, their uses
and properties. To finish we compare the performance, resilience and
quality of POP and our solution.

@@

First we introduce some definitions and notations for mathematical
representation to present our model.

@@

An action is a state operator. It is represented as a tuple
\(a = \langle pre, eff \rangle\) with \(pre\) and \(eff\) being sets of
fluents, respectively the preconditions and the effects of \(a\). An
action can be used only in a state that verifies its preconditions. We
note \(s \models a \Leftrightarrow pre(a) \subset s\) the verification
of an action \(a\) by a state \(s\).

@@

An action \(a\) can be functionally applied to a state \(s\) as follows
: \[a:= \substack{ \left\{ s \models a \middle| s \in S\right\} \to S\\
    a(s) \mapsto s + eff(a)}\] with \(S\) being the set of all states.
This means that an action adds all its effect to the state of the world
on application (by removing complementary fluents if needed)

We distinguish between two specific kinds of actions : actions with no
preconditions are synonymous to a state and those with empty effect are
called a goal.

@@

We define a partial plan satisfaction problem as a tuple noted
\(P = \langle A, I, G, p \rangle\) with :

@@

A \emph{partial plan} is a tuple \(p = \langle A_p, L\rangle\) where
\(A_p\) is a set of steps (actions) and \(L\) is a set of causal links
of the form \(a_i \xrightarrow{f} a_j\), such that
\(\{ a_i, a_j \} \subset A_p \land f \in eff(a_i) \cap pre(a_j)\) or
said otherwise, this causal link means that the fluent \(f\) is provided
by an effect of \(a_i\) to a precondition of \(a_j\). We include the
ordering constraints of PSP in the causal links. An ordering constraint
is noted \(a_i \to a_j\) and means that the plan means \(a_i\) as a step
that is prior to \(a_j\) without specific cause (usually because of
threat resolution). We note the order relation over \(A\) for an action
\(a_i\) that is prior to the action \(a_j\) following all the causal
links and ordering constraints \(a_i \succ a_j\) .

@@

\begin{figure}[htbp]
\centering
\includegraphics{graphics/legend.pdf}
\caption{Global legend for how partial plans are represented in this
paper\label{fig:legend}}
\end{figure}

In this figure the fluent \(5\) is a precondition of the action \(a\)
and the fluent \(6\) an effect. The causal link is providing the fluent
\(7\) (only as a representation example or it would be a liar link). The
dotted line is an ordering constraint that would mean that \(a\) should
be placed before \(a\) (for representation still or it would be a
cycle). \(I\) and \(G\) are respectively the initial and goal step.

@@

A subgoal \(s\) is a precondition of an action \(a_s \in A_p\) with
\(s \in pre(a_s)\) that isn't satisfied by any causal link. We can note
a subgoal as :
\[a_i \xrightarrow{s} a_s \notin L \mid \{ a_i, a_s \} \subset A_p \]

@@

In other words, the action has a possible complementary effect that can
be inserted between two actions needing this fluent while being
consistent with the ordering constraint in \(L\).

\subsubsection{Resolvers}\label{resolvers}

Resolvers are a set of actions and causal links
\(r = \langle A_r , L_r \rangle\) that fixes flaws. A resolver for a
subgoal is an action \(a_r \in A\) that has \(s\) as an effect
\(s \in eff(a_r)\) inserted along with a causal link noted
\(a_r \xrightarrow{s} a_s\). The usual resolvers for a threat are either
\(a_t \to a_i\) or \(a_j \to a_t\) which are called respectively
promotion and demotion links. Another resolver is called a white knight
that is an action \(a_k\) that reestablishes \(t\) after \(a_t\) along
with an ordering constraint \(a_t \to a_k\) and a causal link
\(a_k \xrightarrow{t} a_j\).

\subsubsection{Solution}\label{solution}

The solution of a PSP problem is a valid partial plan that respect the
specification of said problem (only using actions in \(A\) and having
the correct initial and goal step).

@@

A partial plan is consistent if it contains no ordering cycles. That
means that the directed graph formed by step as vertices and causal
links as edges isn't cyclical. This is important to guarantee the
soundness of the algorithm.

@@

We can instantiate one or several flat plans from a partial plan. A flat
plan is an ordered sequence of actions \(\pi = [ a_1, a_2 \ldots a_n]\)
that acts like a pseudo action
\(\pi = \langle pre_\pi, eff_\pi \rangle\) and can be applied to a state
\(s\) using functional composition operation
\(\pi := \bigcirc_{i=1}^n a_n\).

@@

\subsection{Description}\label{description}

@@

\Function{pop}{Queue of Flaws $agenda$, Problem $P$}
\State \Call{populate}{$agenda$, $P$} \Comment{Only on first call}
\If{$agenda = \emptyset$} \State \Return Success
\Comment{Stop all recursion} \EndIf
    \State Flaw \(f\gets agenda.pop\)
\Comment{First element of the queue} \State Resolvers \(R \gets\)
\Call{resolvers}{$f$, $P$} \Comment{Ordered list of resolvers to try}
\ForAll{$r \in R$} \Comment{Choice operator}
\State \Call{apply}{$r$, $P.p$} \If{\Call{consistent}{$P.p$}}
\State \Call{pop}{$agenda \cup$ \protect\Call{relatedFlaws}{$f$}, $P$}
\Else 
            \State \Call{revert}{$r$, $P.p$} \EndIf
    \EndFor
    \State \Return Failure \Comment{Return to last choice of resolver}
\EndFunction

@@

Algorithm 1 presents the base algorithm for a planer in the plan space.
POP implementation uses an agenda of flaws that is efficiently updated
after each refinement of the plan. A flaw is selected for resolution and
we use a non deterministic choice operator to pick a resolver for the
flaw. The resolver is inserted in the plan and we recursively call the
algorithm on the new plan. On failure we return to the last non
deterministic choice to pick another resolver. The algorithm ends when
the agenda is empty or when there is no more resolver to pick for a
given flaw.

\subsection{Limitations}\label{limitations}

This standard way of doing have seen multiple improvements over
expressiveness like with UCPOP {[}7{]}, hierarchical task network to add
more user control over sub-plans {[}8{]}, cognition with defeasible
reasoning {[}9{]}, or speed with multiple ways to implement the popular
fast forward method from state planning {[}10{]}. However, all these
variants do not treat the problem of online planning, resilience and
soft solving.

Some other closer works like {[}4{]} treats the problem of online
planning by removing big chunks of the partial plan by identifying
incorrect trees in the plan. This along with a heuristic of choice of
unrefinemment strategies causes a heavy replanning of the problem even
if only one action needs removal. This also lead to an exponential
number of plan to consider. This is a big problem when trying to adapt a
plan with minimal changes due to replanning.

Indeed, all these problems can affect POP's performance and quality as
they can interfere with POP's inner working when the algorithm is able
to give an answer at all.

Before continuing, we present a simple example of classical POP
execution with the following problem. We have an initial state
\(I = \langle \emptyset , \{ 1, 2 \} \rangle\) and a goal
\(G = \langle \{ 3, 4, -5, 6 \}, \emptyset \rangle\) encoded as dummy
steps. We also introduce actions that are not steps yet but that are
provided by \(A\). The actions \(a \langle \{1\}, \{3,5\} \rangle\),
\(b \langle \{2\}, \{4\} \rangle\) and
\(c \langle \{5\}, \{6\} \rangle\) are normal actions that are useful to
achieve the goal. The actions \(n \langle \{-8, 7\}, \{-7, 8\} \rangle\)
and \(l \langle \{-7, 8\}, \{-8, 7\} \rangle\) are looping actions that
are meant to cause cycles. The action
\(t \langle \{4\}, \{-5\} \rangle\) is meant to be threatening to the
plan's integrity and will generate threats. We introduce
\(u \langle \{4\}, \emptyset \rangle\) and
\(v \langle \{4\}, \{4\} \rangle\) as toxic actions,
\(w \langle \{9\}, \{3\} \rangle\) as a dead-end action and
\(x \langle \{7\}, \{-2,2\} \rangle\) as a contradictory action. These
notions will be defined \protect\hyperlink{defects}{later}.

@@

This example has been crafted to illustrate problems with standard POP
implementations. We give a possible resulting plan of standard POP in
figure \ref{fig:pop}. We can see some issues as for how the plan has
been built. The action \(v\) is being used even if it is useless since
\(b\) already provided fluent \(4\). We can also notice that despite
being contradictory the action \(x\) raised no alarm. As for ordering
constraints, we can clearly see that the link \(a \to t\) is redundant
with the path \(a \xrightarrow{5} c \to t\) that already puts that
constraint by transitivity. Also, some problems arise during execution
with the selection of \(w\) that causes a dead-end.

Of course the flaw selection mechanism of certain variant can prevent
that to happen in that case. But often flaw selection mechanisms are
more speed oriented and will do little if a toxic action seems to fit
better than a more coherent but complex one .

All these issues are caused by what we call \emph{defects} as they are
not regular PSP flaws but they still cause problems with execution and
results. We will address these defects and propose a way to fix them in
\protect\hyperlink{defects}{the next section}.

@@

\Function{properPlan}{Goal $g$, Actions $A$} \State Partial Plan
\(p \gets \emptyset\) \State Actions \(relevants \gets\)
\Call{satisfy}{$g$, $A$, $p$}
\Comment{Satisfy given goal with all necessary actions and causal links}
\State Queue of Actions \(open \gets relevants\)
\While{$open \neq \emptyset$} \State Action \(a\gets open.pop\)
\State Actions \(candidates \gets\) \Call{satisfy}{$a$, $A$, $p$}
\ForAll{$candidate \in candidates$} \If{$candidate \notin relevants$}
\State \(open.push(candidate)\) \EndIf
        \EndFor
    \EndWhile
\EndFunction

@@

As in online planning goals can be known in advance, we propose a new
mechanism that generates proper plans for goals. We take advantage of
the fact that this step can be done offline to improve the performance
of online planning. This offline execution prevents us to access the
details of the initial state of the world as it will be defined at
runtime. We define for that the concept of \emph{participating action}.
An action \(a \in A\) participates in a goal \(G\) if and only if \(a\)
has an effect \(f\) that is needed to accomplish \(G\) or that is needed
to accomplish another participating action's preconditions. A proper
plan is a partial plan that contains all participating actions as steps
and causal links that bind them with the step they are participating in.
This proper plan is independent from the initial step because we might
not have the initial step at the time of the proper plan generation.

@@

We define the concept of \emph{participating action}. An action
\(a \in A\) participates in a goal \(G\) if and only if \(a\) has an
effect \(f\) that is needed to accomplish \(G\) or to accomplish another
participating action's preconditions. A proper plan is a partial plan
that contains all participating actions as steps and causal links that
bind them with the step they are participating in. Proper plan
generation, detailed in Algorithm 2, recursively chooses all
participating actions. It simply populate the proper plan with a quick
and incomplete backward chaining.

The partial plan doesn't have an initial state (because of its offline
nature). This auxiliary algorithm is therefore used as a caching
mechanism for online planning. The algorithm starts to populate the
proper plan with a quick and incomplete backward chaining.

Once applied on the example of figure \textbf{???}, proper plan
generation returns the partial plan presented in figure
\ref{fig:proper}. This partial plan doesn't have initial state because
of its offline nature. It also shows several cycles and obvious
problems. However, it has all the steps of the correct final plan.

This algorithm helps the POP algorithm by prefetching all the actions
subgoals might need but also need to be cleaned first to be helpful.

@@

\Function{clean}{Problem $P$}
\State \Call{assert}{$pre(P.I) = \emptyset$}
\State \Call{assert}{$eff(P.G) = \emptyset$}
\State \(P.p.A_p \gets P.p.A_p \cup \{P.I, P.G\}\)
\State \(defects \gets\) \Call{findDefects}{$P$}
\State \Call{fix}{$defects$, $P$} \EndFunction

\Function{fix}{Defects $defects$, Problem $P$}
\ForAll{Defect $d \in defects$} \State \(defect.\)\Call{fix}{}
\Comment{Fixing is proper to the defect} \EndFor
\EndFunction

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={SODA POP: SOft solving and Defect Aware Partial Ordering Planning},
            pdfauthor={Paper ID\#232},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{listings}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{SODA POP: SOft solving and Defect Aware Partial Ordering Planning}
\author{Paper ID\#232}
\date{}
\usepackage{xcolor, listings}
\usepackage{latex/aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}

@@

%\setlength{\belowcaptionskip}{-2pt}
\setlength{\abovecaptionskip}{-12pt}

@@

\begin{document}
\maketitle
\begin{abstract}
As of recent years, automated planning domain has mainly focused on
performances and advances in state space planning to improve
scalability. That orientation shadows other less efficient ways of doing
like Partial Ordering Planning (POP) that has the advantage to be much
more flexible. This approach generates plans that can be refined easily
and can be used to achieve greater resilience, repairing capabilities
and soft resolution that applications based on online plan recognition
or decision-making can benefit from. This paper presents a set of
algorithms, named solving and Defect Aware Partial Ordering Planning
(SODA POP), that aims at targeting these objectives and maintain a high
plan quality. Our algorithms create proper offline plans for goals and
use an effective defect fixing algorithm to repair input plans even if
these plans are corrupted. SODA POP can also use healer actions and
links in order to always return a valid plan even in cases where the
problem is impossible to solve using problem derivation. Some relevant
properties of these algorithms are analyzed in this article and
experimental results show interesting performances for online planning
and repairing.
\end{abstract}

@@

For some time, Partial Order Planning (POP) has been the most popular
approach to planning resolution. This kind of algorithms are based on
\emph{least commitment strategy} on plan step ordering that allow
actions to be flexibly interleaved during execution {[}1{]}. Thus, the
way the search is made using flexible partial plan as a search space
allowed for more versatility for a wide variety of uses. As of more
recent years, new state space search approach and heuristics {[}2{]}
have been demonstrated to be more efficient than POP planners due to the
simplicity and relevance of state representations opposed to partial
plans {[}3{]}. This have made the search of performance the main axis of
research for planning.

While this goal is justified, it shadows other problems that some
practical applications cause. For example, the flexibility of Plan Space
Planning (PSP) is an obvious advantage for applications needing online
planning: plans can be repaired on the fly as new information and goals
are considered. The idea of using POP for online planning and repairing
plans instead of replanning from scratch is not new {[}4{]}, but has
never been paired with the resilience that some other cognitive
applications may need, especially when dealing with sensors noise in
input data. This resilience makes fixing errors easier than with an
external algorithm as the plan logic allows for context driven decision
on the way to repair the issues. For example, if an action becomes
unrelevant or incoherent, the flaw in the partial plan makes the issue
explicit and therefore easier to fix. Applications might sometimes
provide plans that can contain errors and imperfections that will
decrease significantly the efficiency of the computation and the quality
of the output plan. Additionnaly, these plans may become totally
unsolvable. This problem is to our knowledge not treated in planning of
all forms (state planning, PSP, and even constraint satisfaction
planning) as usually the aim is to find a solution relative to the
original plan (which makes sense). However, as we proceed a mechanism of
\emph{problem derivation} may be required. This will allow soft solving
of any problem regardless of its resolvability.

One of the applications that needs these improvements is plan
recognition with the particular use of off-the-shelf planners to infer
the pursued goal of an agent where online planning and resilience is
particularly important {[}5{]}. This method adds dummy actions that need
to be satisfied by modifying the goal, to ensure that the observed plan
is selected by the planner. A system that is able to repair plan will
ease such an application. Another application is decision-making in
dynamic environments. Indeed, having a plan that details all steps and
explicits the ones that are not possible, can help decide on the plan of
actions to take. These problems call for new ways to improve the answer
of a planner. These improvements must provide relevant plan information
pointing out exactly what needs to be done in order to make a planning
problem solvable, even in the case of obviously inconsistent input data.
This paper aims to solve this while preserving the advantages of PSP
algorithms (flexibility, easy fixing of plans, soundness and
completeness). We propose a new set of auxiliary algorithms that
combined forms our SOft solving and Defect Aware Partial Ordering
Planning (SODA POP) algorithm that targets those issues and allows
making POP algorithm more resilient, efficient and relevant. This is
done by pre-emptively computing proper plans for goals, by solving new
kinds of defects that input plans may exhibit, and by healing
compromised plan by deriving the initial problem with forged actions to
allow resolution. To explain this work we first introduce a few notions,
notations and specificities of existing POP algorithms. Then we present
and illustrate their limitations, categorizing the different defects
arising from the resilience problem and explaining our auxiliary
algorithms, their uses and properties. We compare the performance,
resilience and quality of POP and our solution.

@@

First, we introduce some definitions and notations for mathematical
representation to present our model.

@@

An action is a state operator. It is represented as a tuple
\(a = \langle pre, eff \rangle\) with \(pre\) and \(eff\) being sets of
fluents, respectively the preconditions and the effects of \(a\). An
action can be used only in a state that verifies its preconditions. We
note \(s \models a \Leftrightarrow pre(a) \subseteq s\) the verification
of an action \(a\) by a state \(s\).

@@

An action \(a\) can be functionally applied to a state \(s\) as follows:
\[a:= \substack{ \left\{ s \models a \middle| s \in S\right\} \to S\\
    a(s) \mapsto s + eff(a)}\] with \(S\) being the set of all states.
This means that an action adds all its effects to the state of the world
on application (by removing complementary fluents if needed). We
distinguish between two specific kinds of actions: actions with no
preconditions are synonymous to a state and those with empty effect are
called a goal.

@@

We define a partial plan satisfaction problem as a tuple noted
\(P = \langle A, I, G, p \rangle\) with:

@@

A \emph{partial plan} is a tuple \(p = \langle A_p, L\rangle\) where
\(A_p\) is a set of steps (actions) and \(L\) is a set of causal links
of the form \(a_i \xrightarrow{f} a_j\), such that
\(\{ a_i, a_j \} \subseteq A_p \land f \in eff(a_i) \cap pre(a_j)\) or
said otherwise, this causal link means that the fluent \(f\) is provided
by an effect of \(a_i\) to a precondition of \(a_j\). We include the
ordering constraints of PSP in the causal links. An ordering constraint
is noted \(a_i \to a_j\) and means that the plan means \(a_i\) as a step
that is prior to \(a_j\) without specific cause (usually because of
threat resolution). We note the order relation over \(A\) for an action
\(a_i\) that is prior to the action \(a_j\) following all the causal
links and ordering constraints \(a_i \succ a_j\). We can consider it as
a graph formed by step as vertices and causal links as edges.

@@

\begin{figure}[htbp]
\centering
\includegraphics{graphics/legend.pdf}
\caption{Global legend for partial plans\label{fig:legend}}
\end{figure}

In this figure the fluent \(6\) is a precondition of the action \(a\)
and also an effect. The causal link is providing the fluent \(6\). The
dotted line is an ordering constraint specifying which action should be
placed before another. \(I\) and \(G\) are respectively the initial and
goal step.

@@

A subgoal \(s\) is a precondition of an action \(a_s \in A_p\) with
\(s \in pre(a_s)\) that isn't satisfied by any causal link. We can note
a subgoal as:
\[a_i \xrightarrow{s} a_s \notin L \mid \{ a_i, a_s \} \subseteq A_p \]

@@

In other words, the action has a complementary effect that can be
inserted between two actions needing this fluent while being consistent
with the ordering constraint in \(L\).

\begin{definition}[Resolvers]

Resolvers are a set of actions and causal links
\(r = \langle A_r , L_r \rangle\) that fixes flaws.

\begin{itemize}
\tightlist
\item
  A resolver for a subgoal is an action \(a_r \in A\) that has \(s\) as
  an effect \(s \in eff(a_r)\) inserted along with a causal link noted
  \(a_r \xrightarrow{s} a_s\).
\item
  The usual resolvers for a threat are either \(a_t \to a_i\) or
  \(a_j \to a_t\) which are called respectively promotion and demotion
  links.
\end{itemize}

\begin{definition}[Solution]

The solution of a PSP problem is a valid partial plan that respects the
specification of said problem (only using actions in \(A\) and having
the correct initial and goal step).

@@

A partial plan is consistent if it contains no ordering cycles. The
directed graph formed by step as vertices and causal links as edges
isn't cyclical. This is important to guarantee the soundness.

@@

We can instantiate one or several flat plans from a partial plan. A flat
plan is an ordered sequence of actions \(\pi = [ a_1, a_2 \ldots a_n]\)
that acts like a action \(\pi = \langle pre_\pi, eff_\pi \rangle\) and
can be applied to a state \(s\) using functional composition operation
\(\pi := \bigcirc_{i=1}^n a_n\).

@@

@@

\footnotesize
\Function{pop}{Queue of Flaws $agenda$, Problem $P$}
\State \Call{populate}{$agenda$, $P$} \Comment{Only on first call}
\If{$agenda = \emptyset$} \State \Return Success
\Comment{Stop all recursion} \EndIf
    \State Flaw \(f\gets agenda.popFromQueue\)
\Comment{Stop all recursion} \Comment{First element of the queue}
\State Resolvers \(R \gets\) \Call{resolvers}{$f$, $P$}
\State \Comment{Ordered resolvers to try} \ForAll{$r \in R$}
\Comment{Non deterministic choice operator}
\State \Call{apply}{$r$, $P.p$} \Comment{Apply resolver to partial plan}
\If{\Call{consistent}{$P.p$}} \Comment{$P.p$ is the partial plan}
\State \Call{pop}{$agenda \cup$ \protect\Call{relatedFlaws}{$f$}, $P$}
\State \Comment{Finding new flaws introduced by the resolver} \Else 
            \State \Call{revert}{$r$, $P.p$}
\Comment{Undo resolver insertion} \EndIf
    \EndFor
    \State \Return Failure
\State \Comment{Revert to last non deterministic choice of resolver}
\EndFunction

@@

Algorithm 1 presents the base algorithm for a planer in the plan space.
POP implementation uses an agenda of flaws that is efficiently updated
after each refinement of the plan. A flaw is selected for resolution and
we use a non deterministic choice operator to pick a resolver for the
flaw. The resolver is inserted in the plan and we recursively call the
algorithm on the new plan. On failure, we return to the last non
deterministic choice to pick another resolver. The algorithm ends when
the agenda is empty or when there is no more resolver to pick for a
given flaw. Before continuing, we present a simple example of classical
POP execution with the following example of a problem.

\begin{itemize}
\tightlist
\item
  An initial state \(I = \langle \emptyset , \{ 1, 2 \} \rangle\) and a
  goal \(G = \langle \{ 3, 4, -5, 6 \}, \emptyset \rangle\) encoded as
  dummy steps.
\item
  \(a \langle \{1\}, \{3,5\} \rangle\),
  \(b \langle \{2\}, \{4\} \rangle\) and
  \(c \langle \{5\}, \{6\} \rangle\) are simple actions that are useful
  to achieve the goal.
\item
  \(n \langle \{-8, 7\}, \{-7, 8\} \rangle\) and
  \(l \langle \{-7, 8\}, \{-8, 7\} \rangle\) are looping actions that
  are meant to cause cycles.
\item
  \(t \langle \{4\}, \{-5\} \rangle\) is meant to be threatening to the
  plan's integrity and will generate threats.
\item
  \(u \langle \{4\}, \emptyset \rangle\),
  \(v \langle \{4\}, \{4\} \rangle\),
  \(w \langle \{9\}, \{3\} \rangle\),
  \(x \langle \{7\}, \{-2,2\} \rangle\) are actions that will cause
  problems. \protect\hyperlink{defects}{(cf.~Defects section)}.
\end{itemize}

This standard way of doing has seen multiple improvements over
expressiveness like with UCPOP {[}7{]}, Hierarchical Tasks Network to
add more user control over sub-plans {[}8{]}, cognition with defeasible
reasoning {[}9{]}, or speed with multiple ways to implement the popular
fast forward method from state planning {[}10{]}. However, all these
variants do not treat the problem of online planning, resilience and
soft solving. Some other closer works like {[}4{]} treat the problem of
online planning by identifying and removing trees of the partial plan.
This is done with a heuristic that choses an unrefinemment strategies
that causes a heavy replanning of the problem even if only one action
needs removal. This is a significant problem when trying, for example,
to adapt a plan with minimal changes due to replanning. Indeed, all
these problems can affect POP's performance and quality.

@@

This example has been crafted to illustrate problems with standard POP
implementations. We give a possible resulting plan of standard POP in
figure \ref{fig:pop}. There are some issues as for how the plan has been
built. The action \(v\) is being used even if it is useless since \(b\)
already provided fluent \(4\). We can also notice that despite being
contradictory the action \(x\) raised no alarm. As for ordering
constraints, we can clearly see that the link \(a \to t\) is redundant
with the path \(a \xrightarrow{5} c \to t\) that already ensures that
constraint by transitivity. Also, some problems arise during execution
with the selection of \(w\) that causes a dead-end.

Of course the flaw selection mechanism of certain variants can prevent
that to happen in that case. But often flaw selection mechanisms are
more speed oriented and will do little if a toxic action seems to fit
better than a more coherent but complex one. All these issues are caused
by what we call \emph{defects} as they are not regular PSP flaws but
they still cause problems with execution and results. We will address
these defects and propose a way to fix them in
\protect\hyperlink{defects}{the next section}.

@@

\footnotesize
\Function{properPlan}{Goal $g$, Actions $A$} \State Partial Plan
\(p \gets \emptyset\) \State Actions \(relevants \gets\)
\Call{satisfy}{$g$, $A$, $p$}
\State \Comment{Satisfy the goal with all participating actions and links}
\State Queue of Actions \(open \gets relevants\)
\While{$open \neq \emptyset$} \State Action \(a\gets open.popFromQueue\)
\State Actions \(candidates \gets\) \Call{satisfy}{$a$, $A$, $p$}
\ForAll{$candidate \in candidates$} \If{$candidate \notin relevants$}
\State \(open.pushToQueue(candidate)\) \EndIf
        \EndFor
    \EndWhile
\EndFunction

@@

@@

As in online planning goals can be known in advance, we propose a new
mechanism that generates proper plans for goals. We take advantage of
the fact that this step can be done offline to improve the performance
of online planning. This offline execution prevents us to access the
details of the initial state of the world as it will be defined at
runtime. We define for that the concept of \emph{participating action}.
An action \(a \in A\) participates in a goal \(G\) if and only if \(a\)
has an effect \(f\) that is needed to accomplish \(G\) or that is needed
to accomplish another participating action's preconditions. A proper
plan is a partial plan that contains all participating actions as steps
and causal links that bind them with the step they are participating in.
This proper plan is independent from the initial step because we might
not have the initial step at the time of the proper plan generation.
Once applied on the previous example, proper plan generation returns the
partial plan presented in figure \ref{fig:proper}. This partial plan
doesn't have initial state because of its offline nature. It also shows
several cycles and obvious problems. However, it has all the steps of
the correct final plan. This algorithm helps the POP algorithm by
prefetching all the actions subgoals might need. However, this result
needs to be cleaned first to be helpful.

@@

\footnotesize
\Function{clean}{Problem $P$}
\State \Call{assert}{$pre(P.I) = \emptyset$}
\State \Call{assert}{$eff(P.G) = \emptyset$}
\State \(P.p.A_p \gets P.p.A_p \cup \{P.I, P.G\}\)
\State \(defects \gets\) \Call{findDefects}{$P$}
\State \Call{fix}{$defects$, $P$} \EndFunction

\Function{fix}{Defects $defects$, Problem $P$}
\ForAll{Defect $d \in defects$} \State \(defect.\)\Call{fix}{$P$}
\Comment{Fixing is proper to the defect} \EndFor
\EndFunction
